Cache Oblivious B-trees

### 2011-12-09 paper.tex
Fixing the insert section. I may add updates to it later.

\Search uses $O(log_B N)$ memory transfers where $N = \Theta(c^{F_k})$ and $c$
is your fanout. Note that $c$ is not treated as a constant since this is a
cache-oblivious data structure. This bound is achieved by searching over the
top and bottom recursive subtrees at each level when traversing the tree. A
key trick is that when you visit a node you only search its largest buffer
(that last buffer in the linked list for that level). The smaller buffers will
be visited eventually as you recurse down the tree. Eventually you will get
down to a single subtree that can fit into $O(1)$ blocks and thus can be
searched with $O(1)$ memory accesses. Since the fanout is proportional to our
block size it will take us $O(log_B N)$ block accesses until we get down to a
subtree and thus, by induction, the worst case runtime is $O(log_B N)$.

*****

\Insert uses $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$
memory transfers. This improvement of the the CO B-tree is achieved because
the shuttle tree's buffers are used to maintain locality of the elements long
enough to amortize the cost of moving them down the tree and possible into
another block. The proof is based on the fact that when we move an item from
one block to another, caused by splitting or buffers being overfilled, at
least $B^{\Theta(1/(loglogB)^2)}$ will move with that item. This is because,
with insert, when a buffer overflows we move all of the elements from that
buffer down the tree, not just the single one that caused the overflow. Thus,
we amortize the cost of a single insert by the number of elements in a node's
largest buffer.

******

>>>>

\Search{} uses $O(log_B N)$ memory transfers where $N = \Theta(c^{F_k})$ and
$c$ is your fanout. Note that $c$ is not treated as a constant since this is a
cache-oblivious data structure. This bound is achieved by searching over the
top and bottom recursive subtrees at each level when traversing the tree. A
key trick is that when you visit a node you only search its largest buffer
(that last buffer in the linked list for that level). The smaller buffers will
be visited eventually as you recurse down the tree. Eventually you will get
down to a single subtree that can fit into $O(1)$ blocks and thus can be
searched with $O(1)$ memory accesses. Since the fanout is proportional to our
block size it will take us $O(log_B N)$ block accesses until we get down to a
subtree and thus, by induction, the worst case runtime is $O(log_B N)$.

% TODO - redo the insert section. Namely, include a discussion on how the
% shuttle tree maintains its structure dynamically. I mention the node
% splitting from pg 85 in the paper but do not include that in the analysis. I
% should basically mention that the cost of dynamically maintaining the
% shuttle tree O(logN / B) is not greater than the PMA log^2 N / B and thus
% the dynamic updates do not change the number of memory tranfers.

\Insert{} uses $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$
memory transfers. This improvement over the CO B-tree is achieved because the
shuttle tree's buffers are used to maintain locality of the elements long
enough to amortize the cost of moving them down the tree and possibly into
another memory block. The proof is based on the fact that when an element is
moved from one block to another it is due to either a leaf node splitting or a
node's buffer overflowing. Recall that when inserting a new element causes a
buffer to overflow, all of the the elements from that buffer move down the
tree, not just the single one that caused the overflow. Thus, the cost of
inserting an element is amortize by the number of elements in a node's largest
buffer. The number of elements in the largest buffer is
$B^{\Theta(1/(loglogB)^2)}$ leading to the $B^{\Theta(1/(loglogB)^2)}$ factor
in the big-Oh notation.
